Meta-analysis

### 2019-10-29 05_MetaAnalysis.Rmd
Added a looped analysis to investigate the best incrop rainfall window to predict grain yield

@@

## Visualise Data

@@

## Yield vs in-season rain

@@

Loop over the dates at the start of the season (to 90 days after planting) to find which date fits the best linear model for grain weight and in-crop rainfall.
```{r rainfall_sum_Sstart}
# add column for sum in-crop rainfall
rain_dat_sum$rainfall_sum <- NA

# source funtion that uses bomrang get_historical to return the sum rainfall between two dates
source("R/AUS_rainfall_function.r")
# loop to download and save all the rainfall sum data
# this first loop Investigate when is the first day in the season for which in-crop rainfall is important

for(j in -20:20){
   if(j == -20){lm_data <- data.frame()}

@@

         rain_dat_sum$rainfall_sum[i] <- 
            AUS_rainfall(StartDate = (as.Date(rain_dat_sum$planting_date)[i] + j),
                         EndDate = (as.Date(rain_dat_sum$planting_date)[i] + 90 ),
                         lat = rain_dat_sum$lat[i],
                         long = rain_dat_sum$lon[i])

@@

      lm_data <- rbind(lm_data,
                    data.frame(days_from_plant = j,
                               estimate = coef(lmod1)[2,1],
                               P_value = coef(lmod1)[2,4],
                               r_squared = lmod1$r.squared,
                               adj_r_squared = lmod1$adj.r.squared,
                               stringsAsFactors = FALSE)
                    )

   } # If it aborts with an error Error in row_start:(row_start + season) : argument of length 0, just re-run the code

plot(log(P_value) ~ days_from_plant, data = lm_data)
abline(v = lm_data[lm_data$P_value == min(lm_data$P_value),"days_from_plant"])
plot(r_squared ~ days_from_plant, data = lm_data)
abline(v = lm_data[lm_data$r_squared == max(lm_data$r_squared),"days_from_plant"])
plot(estimate ~ days_from_plant, data = lm_data)
abline(v = 14, h = lm_data[lm_data$r_squared == max(lm_data$r_squared),"estimate"], col = "grey")
```

@@

# Loop to find the last day in the season for which in crop rainfall is important
for(j in -20:20){
   if(j == -10){lm_data <- data.frame()}

@@

         rain_dat_sum$rainfall_sum[i] <- 
            AUS_rainfall(StartDate = (as.Date(rain_dat_sum$planting_date)[i] + 11),
                         EndDate = (as.Date(rain_dat_sum$planting_date)[i] + 90 + j),
                         lat = rain_dat_sum$lat[i],
                         long = rain_dat_sum$lon[i])

@@

      lm_data <- rbind(lm_data,
                    data.frame(days_from_plant = j,
                               estimate = coef(lmod1)[2,1],
                               P_value = coef(lmod1)[2,4],
                               r_squared = lmod1$r.squared,
                               adj_r_squared = lmod1$adj.r.squared,
                               stringsAsFactors = FALSE)
                    )

@@

plot(P_value ~ days_from_plant, data = lm_data)
abline(v = -3)
abline(h = 0.00315, col = "grey")
plot(r_squared ~ days_from_plant, data = lm_data)
abline(v = -3)
abline(h = 0.562, col = "grey")
plot(estimate ~ days_from_plant, data = lm_data)
abline(v = -3)

@@

# use Bomrang to find the quantaty of rain for the season and do a linear regression

@@

@@

# function to extract date from the year
# format should be Universal time format
source("R/year2day.r")
year.day(season_dates$Planting_date[5])

@@

    scale_x_continuous(
      limits = c(0, 200),
      labels = c(),
      breaks = c(1, 50, 100, 150, 200)
   )

as.character(format(FS_dates, "%b-%d"))

as.Date()

head(season_dates)

@@

## Standardising the type of variance

@@

```{r LSD_trials}
Trials <- PM_MB_means %>%
   filter(fungicide_ai == "tebuconazole",
          row_spacing == 0.75,
          Y_error_type == "lsd (P=0.05)") %>%
   select(trial_ref) %>%
   distinct()
Trials
```

@@

Now to estimate the variance from the standard deviation

@@

@@

```{r}
NATrials <- PM_MB_means %>%
   filter(fungicide_ai == "tebuconazole",
          row_spacing.m == 0.75,
          is.na(Y_error_type)) %>%
   select(trial_ref) %>%
   distinct()
```

@@

@@

length(unique(PM_MB_train$Trial_ref))

# Run a linear model with grain_yield.t.ha as a fixed effect predictor and trial as a random intercept
PM_MB_trainLMER <- lmer(yield_error ~ grain_yield.t.ha +
                           (1 | trial_ref), data = PM_MB_train)
summary(PM_MB_trainLMER)

@@

Just for kicks lets look at the linear regression model as a random effects model where `grain_yield.t.ha` is set as a random slope to the random intercept `trial_ref`.

>>>>

# see Yield vs in-season rain section on the influence of rain and irrigation on grain yield
PM_MB_means$irrigation <- NA
PM_MB_means[PM_MB_means$location == "Hermitage", "irrigation"] <- TRUE
PM_MB_means[PM_MB_means$location != "Hermitage", "irrigation"] <- FALSE

@@

## Explore and visualise Data

@@

### Yield vs in-season rain

@@

#### Best start day for in-crop rainfall

Loop over the dates at the start of the season (to 90 days after planting) to find which date fits the best linear model for grain weight and in-crop rainfall.

```{r rainfall_sum_Sstart}
# add column for sum in-crop rainfall
rain_dat_sum$rainfall_sum <- NA

# source funtion that uses bomrang get_historical to return the sum rainfall between two dates
source("R/AUS_rainfall_function.r")

# loop to download and save all the rainfall sum data
# this first loop Investigate when is the first day in the season for which in-crop rainfall is important

# use Bomrang to find the quantaty of rain for the season and do a linear regression
for(k in -19:20){
for(j in -9:30){

   lm_rain <- data.frame(start_day = rep(-20:40, times = length(-30:30)),
              end_day = rep(-30:30, each = length(-20:40)),
              lm_pval = NA,
              lm_rsquared = NA,
              lm_adj_rsquared = NA
   )
   dim(lm_rain)

# for using parRapply   
#    cl8 <- makeCluster(getOption("cl.cores", 8))
# clusterExport(cl8, list("rain_dat_sum", "AUS_rainfall", "lm_rain", "get_historical"))

   testlist1 <- apply(lm_rain,1,function(x1){
         for(i in seq_along(rain_dat_sum$trial_ref)){

         rain_dat_sum$rainfall_sum <- AUS_rainfall(StartDate = as.Date(rain_dat_sum$planting_date)[i] + x1[1],
                                                   EndDate = as.Date(rain_dat_sum$planting_date[i]) + 80 + x1[2],
                                                   lat = rain_dat_sum$lat[i],
                                                   long = rain_dat_sum$lon[i])
         }
         lmod1 <- summary(lm(rainfall_sum ~ grain_yield, 
                       data = rain_dat_sum[rain_dat_sum$location != "Hermitage",])) # remove hermitage which is irrigated

         pval <- coef(lmod1)[2,4]
         rsqu <- lmod1$r.squared
         ADrs <- lmod1$adj.r.squared
         return(list(pval,rsqu,ADrs))

         })

head(testlist1)   
lm_rain$lm_pval <- head(bind_rows(testlist1))
   ?bind_rows

for(i in seq_along(testlist1)){
   lm_rain$lm_pval[i] <- testlist1[[i]][[1]]
   lm_rain$lm_rsquared[i] <- testlist1[[i]][[2]]
   lm_rain$lm_adj_rsquared[i] <- testlist1[[i]][[3]]
}
head(lm_rain)

summary(lm_rain$lm_adj_rsquared)

ggplot(lm_rain, aes(x = start_day, y = end_day, z = lm_pval))+
   geom_contour()

   if(k == -19 & j == -19){
      lm_pval <- matrix(nrow = 40,
                        ncol = 40)
      colnames(lm_pval) <- as.character(-19:20) # End date
      row.names(lm_pval) <- as.character(-9:30) # start date

      lm_square <- matrix(nrow = 40,
                          ncol = 40)
      colnames(lm_square) <- as.character(-19:20) # End date
      row.names(lm_square) <- as.character(-9:30) # start date
      }

@@

         rain_dat_sum$rainfall_sum[i] <- 
            AUS_rainfall(StartDate = (as.Date(rain_dat_sum$planting_date)[i] + j),
                         EndDate = (as.Date(rain_dat_sum$planting_date)[i] + 80 + k ),
                         lat = rain_dat_sum$lat[i],
                         long = rain_dat_sum$lon[i])

@@

      lm_pval[j,k] <- coef(lmod1)[2,4]
      lm_square[j,k] <- lmod1$r.squared

   }
   }# If it aborts with an error Error in row_start:(row_start + season) : argument of length 0, just re-run the code

image(1:ncol(lm_pval), 1:nrow(lm_pval), t(lm_pval), col = terrain.colors(60), axes = TRUE)

image(1:ncol(lm_square), 1:nrow(lm_square), t(lm_square), col = terrain.colors(60), axes = TRUE)

ggplot(lm_pval, aes(x = colnames(lm_pval), y = row.names(lm_pval), z = ))

plot(log(P_value) ~ days_from_plant, data = lm_data)
abline(v = lm_data[lm_data$P_value == min(lm_data$P_value),"days_from_plant"], col = "grey")
text(x = lm_data[lm_data$P_value == min(lm_data$P_value),"days_from_plant"], 
     y = log(max(lm_data$P_value))/0.95, 
     labels = paste("pvalue =",round(min(lm_data$P_value),2)),
     pos = 4)
plot(r_squared ~ days_from_plant, data = lm_data)
abline(v = lm_data[lm_data$r_squared == max(lm_data$r_squared),"days_from_plant"], col = "grey")
text(x = lm_data[lm_data$r_squared == max(lm_data$r_squared),"days_from_plant"], 
     y = max(lm_data$r_squared * 0.75), 
     labels = paste("Rsquared =",round(max(lm_data$r_squared),4)),
     pos = 2)

plot(estimate ~ days_from_plant, data = lm_data)
abline(v = 14, h = lm_data[lm_data$r_squared == max(lm_data$r_squared),"estimate"], col = "grey")
```

#### Best end day for in-crop rainfall

@@

# Loop to find the last day in the season for which in crop rainfall is important
for(j in -20:20){
   if(j == -20){lm_data <- data.frame()}

@@

         rain_dat_sum$rainfall_sum[i] <- 
            AUS_rainfall(StartDate = (as.Date(rain_dat_sum$planting_date)[i] + 14),
                         EndDate = (as.Date(rain_dat_sum$planting_date)[i] + 80 + j),
                         lat = rain_dat_sum$lat[i],
                         long = rain_dat_sum$lon[i])

@@

      lm_data <- rbind(lm_data,
                    data.frame(days_from_plant = j + 80,
                               estimate = coef(lmod1)[2,1],
                               P_value = coef(lmod1)[2,4],
                               r_squared = lmod1$r.squared,
                               adj_r_squared = lmod1$adj.r.squared,
                               stringsAsFactors = FALSE)
                    )

@@

plot(log(P_value) ~ days_from_plant, data = lm_data, xlab = "End of in-crop rain referenced to 90 days after planting")
abline(v = lm_data[lm_data$P_value == min(lm_data$P_value),"days_from_plant"], col = "grey")
text(x = lm_data[lm_data$P_value == min(lm_data$P_value),"days_from_plant"], 
     y = max(log(lm_data$P_value * 0.75)), 
     labels = paste("pvalue =",round(min(lm_data$P_value),4)),
     pos = 4)

plot(r_squared ~ days_from_plant, data = lm_data, xlab = "End of in-crop rain referenced to 90 days after planting")
abline(v = lm_data[lm_data$r_squared == max(lm_data$r_squared),"days_from_plant"], col = "grey")
text(x = lm_data[lm_data$r_squared == max(lm_data$r_squared),"days_from_plant"], 
     y = max(lm_data$r_squared * 0.75), 
     labels = paste("Rsquared =",round(max(lm_data$r_squared),4)),
     pos = 2)

plot(estimate ~ days_from_plant, data = lm_data)
abline(v = -3 ,h = lm_data[lm_data$r_squared == max(lm_data$r_squared),"estimate"], col = "grey")

@@

@@

### Season range and first incidence

@@

# function to extract date from the year
# format should be Universal time format
source("R/year2day.r")

@@

@@

## Standardising sample variance

@@

### Converting least-squares tp sample variance

```{r all_LSD_trials}
Trials <- PM_MB_means %>%
   filter(
      #fungicide_ai == "tebuconazole",
      #    row_spacing == 0.75,
          Y_error_type == "lsd (P=0.05)") %>%
   select(trial_ref) %>%
   distinct()
Trials
```

@@

### Converting standard deviation to sample variance
Now to estimate the variance from the standard deviation

@@

### Estimating variance for trials without reported error

@@

```{r}
NATrials <- PM_MB_means %>%
   filter(fungicide_ai == "tebuconazole",
          row_spacing == 0.75,
          is.na(Y_error_type)) %>%
   select(trial_ref) %>%
   distinct()
```

@@

#### Mixed effect model to estimate variance

@@

length(unique(PM_MB_train$trial_ref))

# Run a linear model with grain_yield.t.ha as a fixed effect predictor and trial as a random intercept
PM_MB_trainLMER <- lmer(yield_error ~ grain_yield.t.ha. +
                           (1 | trial_ref), data = PM_MB_train)
summary(PM_MB_trainLMER)

@@

#### Random effects model to predict trial variance
Just for kicks lets look at the linear regression model as a random effects model where `grain_yield.t.ha` is set as a random slope to the random intercept `trial_ref`.
